---
title: "cbdb"
author: "Julia Ulziisaikhan"
date: "1/26/2022"
output: html_document
---
```{r, warning=F, message=F}
library(tidyverse)
```

# Processing data and creating the cbdb_ddm_bdrc.csv file

## CBDB

Loading in CBDB

```{r}
biog_main <- read.csv("C:/Users/Julia/Documents/GitHub/tibetan-longevity/biog_main/cbdb_persons_data.csv")
# data on ethnicities
eth <- 
  read.csv("C:/Users/Julia/Documents/GitHub/tibetan-longevity/biog_main/cbdb_ethnicity_codes.csv") %>% 
  select(c_ethnicity_code, c_name, c_group_code)
```

```{r}
head(eth)
```


Processing data. Excluding persons with empty birthyears. Joining the table with the ethnicity codes table, and only including 0 and 1 - one of which I believe is Han and the other is unknown or NA.

```{r}
cbdb <- biog_main %>% 
  filter(!is.na(c_birthyear) & (c_birthyear != 0) & (c_deathyear != 0)) %>%
  mutate(lifespan = c_deathyear-c_birthyear) %>% 
  relocate(c('c_birthyear','c_deathyear', 'lifespan'), .after='c_personid')  %>%
  arrange(-c_birthyear) %>%
  select(c_personid, c_birthyear, c_deathyear, lifespan, c_female, c_ethnicity_code, c_name_chn) %>% 
  left_join(eth, by = "c_ethnicity_code") %>%
  mutate(c_personid = as.character(c_personid)) %>%
  filter(c_ethnicity_code == 0 | c_ethnicity_code == 1) %>%
  select(-c_ethnicity_code, -c_name, -c_group_code)

cbdb$data <- "cbdb"

head(cbdb)
```

Note: Not sure why the column `c_death_age` provided by CBDB is always 1 year higher than our own calculated lifespan column (where we = c_deathyear - c_birthyear). 

## DDM and BDRC

We import data from Chinese monks (DDM) and Tibetan figures (BDRC) already processed in `ddm_bdrc_analysis.Rmd`

```{r}
ddm_bdrc <- 
  read.csv("C:/Users/ulzii/OneDrive/Documents/spring22/tuttle/biog_main/bdrc_and_ddm_data.csv")
head(ddm_bdrc)
```

Columns are not the same between ddm_bdrc and cbdb, so we must change that before we can rbind() or dplyr::bind_rows()

```{r}
names(ddm_bdrc)
```

```{r}
names(cbdb)
```
Changing column names so they are consistent.

```{r}
names(cbdb) <- c("id", "bdate", "ddate", "lifespan", "female", "name", "data")
cols <- names(cbdb)[names(cbdb) %in% names(ddm_bdrc) == F]
ddm_bdrc[,cols] <- NA
```

```{r}
df <- 
  dplyr::bind_rows(cbdb, ddm_bdrc) %>%
  select(id, data, bdate, ddate, lifespan, name, seats, female) %>%
  filter(bdate >= 500) %>%
  arrange(bdate)
write.csv(df, "C:/Users/ulzii/OneDrive/Documents/spring22/tuttle/biog_main/cbdb_ddm_bdrc.csv", row.names=F)
```

```{r}
glimpse(df)
```

```{r}
df <- read.csv("C:/Users/ulzii/OneDrive/Documents/spring22/tuttle/biog_main/cbdb_ddm_bdrc.csv")
glimpse(df)
```

# Exploratory Data Analysis

```{r}
biog_main %>% filter(c_name == "tu bo")
```


```{r}
table(df$data) %>% 
  data.frame() %>% 
  mutate(prop = (Freq / sum(Freq))%>% round(3) * 100 )
```
```{r}
df %>%
  filter(!is.na(female)) %>%
  select(female) %>%
  table() %>% 
  data.frame() %>% 
  mutate(prop = (Freq / sum(Freq))%>% round(3) * 100 )
```

```{r}
df %>%
  group_by(ethnicity_code) %>%
  summarise(n = n()) %>%
  arrange(-n) %>%
  left_join(., eth, by = c("ethnicity_code"= "c_ethnicity_code")) %>%
  select(-c_group_code)
```

```{r}
glimpse(df)
```

```{r}
base <- 
  ggplot(data=df, aes(x=ddate, y=lifespan, colour=data, fill=data)) + 
  geom_point(alpha=0.1) 

base +
  scale_x_continuous(breaks = seq(1000, 2000, 100), limits=c(1000,2000), name = "Year of Death", expand = c(0, 0)) +
  scale_y_continuous(breaks = seq(10, 100, 10), limits=c(0,100), name = "Lifespan (Years)", expand = c(0, 0)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.background = element_rect(fill = "grey90"),
        axis.text.x = element_text(size=9, angle=30, vjust = 1.25, hjust=1.1),
        axis.line = element_line(colour = "black", size = 0.3, linetype = "solid"),
        panel.background = element_rect(fill = "white", color='black'),
        panel.grid.major = element_line(color = 'grey85', size=0.5),
        panel.grid.minor = element_line(color = 'grey85', size=0.5)
        ,        legend.position="bottom"
        )
```

```{r}
df %>% filter(ethnicity_code != 0 & ethnicity_code != 1 & data == "cbdb") %>% group_by(ethnicity_name) %>% summarise(n=n()) %>% arrange(-n)
```


```{r}
new <- 
  df %>%
  group_by(data) %>%
  sample_n(2328)

base2 <- 
  ggplot(data=df %>% filter(data=="cbdb"), aes(x=ddate, y=lifespan, colour=data, fill=data)) + 
  geom_point(alpha=0.2) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.background = element_rect(fill = "grey90"),
        axis.text.x = element_text(size=9, angle=30, vjust = 1.25, hjust=1.1),
        axis.line = element_line(colour = "black", size = 0.3, linetype = "solid"),
        panel.background = element_rect(fill = "white", color='black'),
        panel.grid.major = element_line(color = 'grey85', size=0.5),
        panel.grid.minor = element_line(color = 'grey85', size=0.5)
        ,        legend.position="bottom"
        )

base2 + scale_x_continuous(limits=c(1000,2000)) + geom_smooth()
```


```{r}
ggplot(data=(df %>% filter(data=="cbdb")), aes(x=ddate, y=lifespan))

df
```


```{r}
ggplot(data=cbdb, aes(x=ddate, y=lifespan)) + 
  geom_point(alpha=0.1) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.background = element_rect(fill = "grey90"),
        axis.text.x = element_text(size=9, angle=30, vjust = 1.25, hjust=1.1),
        axis.line = element_line(colour = "black", size = 0.3, linetype = "solid"),
        panel.background = element_rect(fill = "white", color='black'),
        panel.grid.major = element_line(color = 'grey85', size=0.5),
        panel.grid.minor = element_line(color = 'grey85', size=0.5)
        ,        legend.position="bottom"
        ) +
  geom_smooth()
```

11/18/2022

```{r}
df_w <- 
  read.csv("C:/Users/Julia/Documents/GitHub/tibetan-longevity/wikidata_analysis/data/wikidata_raw.csv") %>% 
  mutate(lifespan = yod - yob) %>%
  filter(!is.na(lifespan)) %>% 
  filter(lifespan <= 115 & lifespan > 0) %>%
  group_by(occupationLabel) %>%
  # #comment out the below lines if you want all occupations! ctrl+shift+c
  #filter(occupationLabel %in% religious_occupations | occupationLabel=="")%>%
  rename(
    bdate = yob,
    ddate = yod
    )
df_w <- df_w[!duplicated(df_w$itemLabel), ]
head(df_w)
```

```{r}
ggplot(data=df_w, aes(x=ddate, y=lifespan)) + 
  geom_point(alpha=0.2)  +
  scale_x_continuous(breaks = seq(1000, 2000, 10), limits=c(1900,2000), name = "Year of Death", expand = c(0, 0)) +
  scale_y_continuous(breaks = seq(10, 115, 10), limits=c(0,115), name = "Lifespan (Years)", expand = c(0, 0)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.background = element_rect(fill = "grey90"),
        axis.text.x = element_text(size=9, angle=30, vjust = 1.25, hjust=1.1),
        axis.line = element_line(colour = "black", size = 0.3, linetype = "solid"),
        panel.background = element_rect(fill = "white", color='black'),
        panel.grid.major = element_line(color = 'grey85', size=0.5),
        panel.grid.minor = element_line(color = 'grey85', size=0.5)
        ,        legend.position="bottom"
        ) +geom_smooth() + ggtitle("all european figures")
```

