---
title: "ggplot_timeseries"
author: "julia"
date: "2022-10-21"
output: html_document
---
This is the main file

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(ggfortify)
library(changepoint)
library(strucchange)
library(ggpmisc)
```

Loading in data of CBDB, DDM, and BDRC. This table was processed in a different R file.

```{r}
data <- read.csv("C:/Users/Julia/Documents/GitHub/tibetan-longevity/biog_main/cbdb_ddm_bdrc.csv")
head(data)
```

```{r}
df <- 
  filter(data, data=="bdrc") %>% 
  #post 1000
  filter(., ddate > 1000) %>%
  select(ddate) %>% 
  mutate(count = 1) %>% 
  group_by(ddate) %>% 
  summarise(count = n()) %>% 
  rename(year=ddate)
```

We load in WikiData persons data (European-focus). Below I am

```{r}
religious_occupations <- c("Catholic priest", "missionary", "priest", "rabbi", "cardinal", "feudatory", "Christian minister", "religious", "pastor", "nun", "monk", 
                           "Anglican priest", "Latin Catholic priest", "abbess", "Catholic friar", "Catholic theologian", "Christian nun")
religious_occupations

data_w <- 
  read.csv("C:/Users/Julia/Documents/GitHub/tibetan-longevity/wikidata_analysis/data/wikidata_raw.csv") %>% 
  mutate(lifespan = yod - yob) %>%
  filter(!is.na(lifespan)) %>% 
  filter(lifespan <= 115 & lifespan > 0) %>%
  group_by(occupationLabel) %>%
  #comment out the below lines if you want all occupations! ctrl+shift+c
  # filter(occupationLabel %in% religious_occupations | occupationLabel=="")%>%
  rename(
    bdate = yob,
    ddate = yod
    )
data_w <- data_w[!duplicated(data_w$itemLabel), ]
head(data_w)
nrow(data_w)
```


```{r}
df_w <-
  #post 1000
  filter(data_w, ddate > 1000) %>%
  select(ddate) %>% 
  mutate(count = 1) %>% 
  group_by(ddate) %>% 
  summarise(count = n()) %>% 
  rename(year=ddate)

head(df_w)
```


```{r}
years <- 
  seq(range(df_w$year)[1], range(df_w$year)[2]) %>% 
  data.frame() %>% 
  rename(year=".")
```


```{r}
df2 <- left_join(years, df, by = c("year"="year"))%>%mutate_all(funs(replace_na(.,0)))
head(df2)
```

```{r}
df2_w <- left_join(years, df_w, by = c("year"="year"))%>%mutate_all(funs(replace_na(.,0)))
head(df2_w)
```



```{r}
ma <- function(x, n = 5){stats::filter(x, rep(1 / n, n), sides = 2)}
```

# n = 1

## wiki

##data highly skewed to post 1800
```{r}
point <- subset(df2_w, year == 1959)
df3_w<-df2_w %>% mutate(count=ma(count, n=1))

ggplot(data=df3_w, aes(x=year, y=ma(count, n=1), group=1)) +
  # geom_bar(stat='identity') +
  geom_line() +
  scale_y_continuous(breaks = seq(0, 50, 10), limits=c(0,50), name = "Number of Deaths", expand = c(0, 0)) +
  scale_x_continuous(breaks = seq(1000, 2100, 100), limits=c(1000,2020), name = "Year", expand = c(0, 0)) +    ggtitle("Plot of number of reported deaths over time") +
  theme_classic() 
```

let's limit to priests only and see if it normalizes the data imbalance. i like the following occupations:
religious: Catholic priest, missionary, priest, rabbi, cardinal, feudatory, Christian minister, religious, pastor, nun, monk, Anglican priest, Latin Catholic priest, abbess, Catholic friar,Catholic theologian, Christian nun
civil: politician, military personnel, diplomat, military officer, soldier
art: painter, composer, architect, artist, sculptor, conductor, astronomer

```{r}
data_w %>% 
  group_by(occupationLabel) %>% 
  summarise(n=n()) %>% 
  arrange(desc(n))
```




```{r}
point <- subset(df2_w, year == 1959)
df3_w<-df2_w %>% mutate(count=ma(count, n=1))

ggplot(data=df3_w, aes(x=year, y=ma(count, n=5), group=1)) +
  # geom_bar(stat='identity') +
  geom_line() +
  scale_y_continuous(breaks = seq(0, 5, 1), limits=c(0,5), name = "Number of Deaths", expand = c(0, 0)) +
  scale_x_continuous(breaks = seq(1000, 2100, 100), limits=c(1000,2020), name = "Year", expand = c(0, 0)) +    ggtitle("Plot of number of reported deaths over time") +
  theme_classic() 
```

## bdrc

```{r}
point <- subset(df2, year == 1959)

ggplot(data=df2, aes(x=year, y=ma(count, n=1), group=1)) +
  # geom_bar(stat='identity') +
  geom_line() +
  scale_y_continuous(breaks = seq(0, 50, 10), limits=c(0,50), name = "Number of Deaths", expand = c(0, 0)) +
  scale_x_continuous(breaks = seq(1000, 2100, 100), limits=c(1900,2020), name = "Year", expand = c(0, 0)) +    ggtitle("Plot of number of reported deaths over time") +
  theme_classic() + 
 geom_point(data=point, colour="red") +  # this adds a red point
  geom_text(data=point, label="1959", vjust=-.5, colour="red") # this adds a label for the red point
```

# n = 2

```{r}
ex <- data.frame(year = 1, count = 3)
print(ex)
```


```{r}
df3<-df2 %>% mutate(count=ma(count, n=2))
point3 <- data.frame(year = 1958, count = 36.2)

ggplot(data=df3, aes(x=year, y=count, group=1)) +
  geom_line() +
  scale_y_continuous(breaks = seq(0, 40, 10), limits=c(0,40), name = "Number of Deaths", expand = c(0, 0)) +
  scale_x_continuous(breaks = seq(1000, 2100, 100), limits=c(1500,2020), name = "Year", expand = c(0, 0)) + ggtitle("Reported deaths over time") +
  theme_classic()+ 
  geom_point(data=point3, colour="red") +  # this adds a red point
  geom_text(data=point3, label="1959", vjust=-.5, colour="red") # this adds a label for the red point
```



```{r}
bdrc_ts <- ts(df2$count, frequency=1, start=1000)
autoplot(bdrc_ts)
```
```{r}
bdrc_ts %>%
  changepoint:: cpt.meanvar() %>%  # Identify change points
  autoplot()

strucchange::breakpoints(bdrc_ts ~ 1) %>%
  autoplot()
```


